---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# debkeepr: Analysis of Non-Decimal Currencies

[![Travis build status](https://travis-ci.org/jessesadler/debkeepr.svg?branch=master)](https://travis-ci.org/jessesadler/debkeepr)
[![Coverage status](https://codecov.io/gh/jessesadler/debkeepr/branch/master/graph/badge.svg)](https://codecov.io/github/jessesadler/debkeepr?branch=master)

`debkeepr` provides an interface for working with non-decimal currencies that use the tripartite system of pounds, shillings, and pence. The package includes functions to apply arithmetic and financial operations to single or multiple values and to analyze account books that use either [single-entry bookkeeping](https://en.wikipedia.org/wiki/Single-entry_bookkeeping_system) or [double-entry bookkeeping](https://en.wikipedia.org/wiki/Double-entry_bookkeeping_system) with the latter providing the name for `debkeepr`.

## Installation
You can install `debkeepr` from GitHub with [devtools](https://github.com/hadley/devtools):

```{r devtools, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jessesadler/debkeepr")
```

## lsd Class
The pounds, shillings, and pence monetary system complicates even relatively simple arithmetic manipulations, as each unit has to be [normalized](https://en.wikipedia.org/wiki/Arithmetic#Compound_unit_arithmetic) or converted to the correct base. `debkeepr` uses the [vctrs package](https://vctrs.r-lib.org) to integrate non-decimal currencies into standard R practices. `debkeepr` introduces two classes: `deb_lsd` and `deb_decimal`. The `deb_lsd` maintains the tripartite structure of most non-decimal currencies and provides a `bases` attribute to define the bases of the solidus or shillings unit and denarius or pence unit. The `deb_decimal` class represents the same values as `deb_lsd` in decimal form. It too provides a `bases` attribute, as well as a `unit` attribute so that values can be represented in either libra (pounds), solidus, or denarius units.

## Historical Background
The `debkeepr` package uses the nomenclature of [l, s, and d](https://en.wikipedia.org/wiki/%C2%A3sd) to represent pounds, shillings, and pence units. The abbreviations derive from the Latin terms [libra](https://en.wikipedia.org/wiki/French_livre), [solidus](https://en.wikipedia.org/wiki/Solidus_(coin)), and [denarius](https://en.wikipedia.org/wiki/Denarius). The libra was a Roman measurement of weight, while the solidus and denarius were both Roman coins. The denarius was a silver coin from the era of the Republic, in contrast to the golden solidus that was issued in the Late Empire. As the production of silver coins overtook that of gold by the 8th century, a solidus came to represent 12 silver denarii coins, and 240 denarii were — for a time — made from one libra or pound of silver. The custom of counting coins in dozens (solidi) and scores of dozens (librae) spread throughout the Carolingian Empire and became engrained in much of Europe. However, a variety of [other bases](https://en.wikipedia.org/wiki/Non-decimal_currency) for the solidi and denarii units were in use in different regions and at different times. `debkeepr` provides a consistent manner for dealing with any set of bases within a tripartite system through the `bases` attribute of an `lsd` object.

## Overview
- All of the functions in `debkeepr` begin with the prefix `deb_`, which is short for double-entry bookkeeping.
- The nomenclature used throughout the package follows the [original Latin terms](https://en.wikipedia.org/wiki/%C2%A3sd) in using l, s, and d to represent librae, solidi, and denarii respectively and to refer to such values as lsd values. These terms were translated into the various European languages.
	- English: pounds, shillings, pence
	- French: livres, sols or sous, deniers
	- Italian: lire, soldi, denari
	- Dutch: ponden, schellingen, groten or penningen

## Examples
At the heart of `debkeepr` is the need to normalize pounds, shillings, and pence values to specified non-decimal unit bases in the process or making various calculations. Even in the simplest arithmetic operations can be tricky with non-decimal currencies. `debkeepr` extracts away much of the difficulties and in many cases makes it possible to procede as if you are working with normal decimalized data. When working with decimalized data is preferable, the `deb_decimal` class make casting from and to `deb_lsd` easy.

For example, adding together a set of values by hand might result in the non-standard form of £131 62s. 41d. in a currency with the standard bases of 20 shillings per £1 and 12 pence per shilling.

```{r addition}
library(debkeepr)

# Normalize £131 62s. 41d.
x <- deb_lsd(131, 67, 42)
deb_normalize(x)

# Or create a deb_lsd vector and add
(y <- deb_lsd(l = c(15, 32, 18, 54, 12),
             s = c(9, 17, 8, 15, 18),
             d = c(11, 8, 9, 4, 10)))

# Do the addition
sum(y)
```

Other arithmetic operations are also made available by `debkeepr`. If you think of `y` as profits to be divided evenly among three parterns, the calculations can be made quickly.

```{r division}

# Profits for each partner
sum(y) / 3
```

`y` could also be transformed into its decimalized form in any of the three units.

```{r deb-decimal}
# Decimalized libra
deb_as_decimal(y)

# Decimalized solidus
deb_as_decimal(y, unit = "s")

# Decimalized denarius
deb_as_decimal(y, unit = "d")
```
